Supplementary Information

### 2021-06-11 doc/manuscript/ForestGEO_treerings_SI.Rmd
google docs: misc cleanup
citations problem #133
minor fixes to SI

Harvard Forest (HF) is a secondary temperate mixed forest dominated by *Tsuga canadensis* and northern hardwood species including *Quercus rubra*, *Acer rubrum*, and *Betula alleghaniensis*.
Cores were collected within the ForestGEO megaplot, and in the footprint of the original HF eddy-flux tower [@finzi_carbon_2020]. 
Samples were collected following an ecological study design [@dye_comparing_2016], wherein all living or dead trees $\ge$ 10 cm DBH within a **(SIZE)** plot were cored.

@@

In total, 2210 individuals were cored either within the 25 ha ForestGEO plot or a larger, 74 ha plot in which it is embedded. 
The tree-ring data were collected for several studies around 2010 [@samonil_individualbased_2013; @samonil_disturbances_2014; @samonil_uncertainty_2015; @samonil_uncertainty_2016; @vasickova_true_2016; @vasickova_driving_2019; @kaspar_speciesspecific_2021]. 
The sampling strategy was mainly connected with the disturbance history of the forest and evaluation of growth of juvenile trees in gaps. 
Therefore, these data consist mainly of dominant and codominant tree species and relatively low number of young trees.

@@

### Cedar Breaks, Utah, USA
Cores were collected within the Utah Forest Dynamics Plot ForestGEO plot on the Colorado Plateau in Cedar Breaks National Monument [CB; @furniss_reconciling_2017]. 
The plot sits on the edge of cliff, which descends ~750 m to a valley below.
The mature spruce-fir forest is dominated by the species analyzed here (*Abies bifolia*, *Populus tremuloides*, *Picea engelmannii*, *Pinus flexilis*, *Pinus longaeva*, *Picea pungens*, and *Pseudotsuga menziesii*).
The site is largely undisturbed. 
The most recent fire occurred in 1802, and many currently living trees established before then. Living *Pinus longaeva* at the site date back ~1600 years, and deadwood at the site dates back to 0 AD [@lutz_largediameter_2021].

>>>>

Harvard Forest (HF) is a secondary temperate mixed forest dominated by *Tsuga canadensis* and northern hardwood species including *Quercus rubra*, *Acer rubrum*, and *Betula alleghaniensis*.
Cores were collected at two plots within the ForestGEO megaplot, and in the footprint of the original HF eddy-flux tower [@finzi_carbon_2020]. 
Samples were collected following an ecological study design of nested circular plots [@dye_comparing_2016], where sampling included all living or dead trees $\ge$ 10 cm DBH within a 13-m radius and $\ge$ 20 cm within a 20-m radius.

@@

In total, 2210 individuals were cored either within the 25 ha ForestGEO plot or a larger, 74 ha plot in which it is embedded. 
The tree-ring data were collected for several studies around 2010 [@samonil_individualbased_2013; @samonil_disturbances_2014; @samonil_uncertainty_2015; @samonil_uncertainty_2016; @vasickova_true_2016; @vasickova_driving_2019; @kaspar_species_2021]. 
The sampling strategy was mainly connected with the disturbance history of the forest and evaluation of growth of juvenile trees in gaps. 
Therefore, these data consist mainly of dominant and codominant tree species and relatively low number of young trees.

@@

### Cedar Breaks, Utah, USA
Cores were collected within the Utah Forest Dynamics Plot ForestGEO plot on the Colorado Plateau in Cedar Breaks National Monument [CB, @furniss_reconciling_2017]. 
The plot sits on the edge of cliff, which descends ~750 m to a valley below.
The mature spruce-fir forest is dominated by the species analyzed here (*Abies bifolia*, *Populus tremuloides*, *Picea engelmannii*, *Pinus flexilis*, *Pinus longaeva*, *Picea pungens*, and *Pseudotsuga menziesii*).
The site is largely undisturbed. 
The most recent fire occurred in 1802, and many currently living trees established before then. Living *Pinus longaeva* at the site date back ~1600 years, and deadwood at the site dates back to 0 AD [@lutz_largediameter_2021].
